% Acknowledgements

\pdfbookmark[1]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer

%----------------------------------------------------------------------------------------

\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements}

\noindent Many are the people that I feel obliged to mention and that I feel have
had a deep contribution to making me grow, both as a student and as a person.

First, I would like to express my deepest and most sincere gratitude
to Professor \textit{Kazuhisa Ito} for giving me the opportunity to visit
his laboratory in Japan, and for his continuous understanding of the challenges
I had in that environment. I really wish I could do more to express how grateful I am.
I would also like to thank all the members of Professor Ito's laboratory,
with whom I feel I have developed an intense bond and that I will never forget.
In particular I want to thank \textit{Kenji Akiba}, with whom I feel I developed a
deep friendship stronger than any distance or language barrier. I want to thank
also \textit{Renato Moura}, \textit{Ryo Inada}, \textit{Yusuke Minato},
\textit{Yuki Tsuchiya}, \textit{Mine-chan}, \textit{Keita Aoyagi},
\textit{Masaru Masuda}, \textit{Andre Pedroso}, \textit{Akira Tsushima}.
We will meet again for sure, in the future.

I want to thank all the people I had the luck to live together with
during my stay in Japan, including but not limited to:
my brother \textit{Davide}, who I often unfairly treat as a relief valve
for my own frustrations and I cannot understand how he hasn't killed me already.
\textit{Daniele Di Girolamo}, with whom I've had the luck to
live with for five months, and proved to be a wonderful person and friend.
\textit{Angelo Damiani}, who has been a friend for years and, I hope, for the years to come.
\textit{Petko Slavov}, for whom I feel a deep respect and admiration.
\textit{Kaz}, which is one of the funniest Japanese guys I've ever met.
\textit{Jimmy} for being friendly to us, even if we sadly didn't have much
time to spend together.

I need to mention other colleagues and professors that I've had the luck to know
and work with: my dear friend \textit{Alessandro D'Ortenzio} who's accompanied
my years and proved to be one of the most reliable and dearest friends.
\textit{Silvia Montecchia}, which has often been a precious confidant
and helped me through gloomy times. Professor \textit{Costanzo Manes} who
has agreed to be my supervisor even if I'm working in a field that is not properly mine,
and followed me with patience and understanding.
Professor \textit{Pierluigi Beomonte Zobel} for giving me the opportunity of a lifetime, 
letting me visit a far away country and experience a diverse culture.
Professors \textit{Gabriele Di Stefano}, \textit{Serafino Cicerone},
\textit{Daniele Frigioni} for their enthusiastic guidance through the years.

Lastly I want to thank my family: my mother \textit{Patrizia}, my father \textit{Giuseppe}
my brother \textit{Davide}, my sisters \textit{Irene} and \textit{Giulia}, my grandmother \textit{Giulia}
and all the other relatives that I don't have the space to write,
and that I feel lucky to be part of the same family.
They have provided support through these years, which sometimes have not been easy,
but that I hope have paved the way for a brighter future.
I wouldn't be who I am without your love, patience, understanding, empathy and care.
I thank each one of you from the deepest of my heart.

And is from the very bottom of my heart that goes my thought
for those who aren't on this earth any more,
as there isn't a day that goes by that I don't think about them.
I want to dedicate this thought to my late grandmother \textit{Maria Rita},
my late grandfathers \textit{Emilio} and \textit{Ireneo}, my late uncle \textit{Silvio},
my late aunt \textit{Nicoletta} and all the others that were and are no more.
I know for sure that you're by my side today, and you've always been.




\endgroup